### 2015-12-02 ML/hw6/hw.tex
ml homework and some random previously unsubmitted files for paralle_programming and HiWi

\begin{equation}
\begin{align}
 \hat{\mathbf{w}}_{MLE} &= (\hat{\Phi}^T\hat{\Phi})^{-1}\hat{\Phi}^T\hat{\mathbf{z}} \\
	       &= \left( \begin{pmatrix} \Phi^T & \sqrt{\lambda}I_{p \times M}^T \end{pmatrix} \begin{pmatrix} \Phi \\ \sqrt{\lambda}I_{p \times M} \end{pmatrix} \right)^{-1} \begin{pmatrix} \Phi^T & \sqrt{\lambda}I_{p \times M}^T \end{pmatrix} \begin{pmatrix} \mathbf{z} \\ \mathbf{0}_p \end{pmatrix} \\
	       &= \left( \lambda I_{M \times M} + \Phi^T\Phi \right)^{-1} \Phi^T \mathbf{z} \\
	       &= \mathbf{w}_{MLE}^{ridge}
\end{align}
\end{equation}

>>>>

%\begin{equation}
\begin{align}
 \hat{\mathbf{w}}_{MLE} &= (\hat{\Phi}^T\hat{\Phi})^{-1}\hat{\Phi}^T\hat{\mathbf{z}} \\
	       &= \left( \begin{pmatrix} \Phi^T & \sqrt{\lambda}I_{p \times M}^T \end{pmatrix} \begin{pmatrix} \Phi \\ \sqrt{\lambda}I_{p \times M} \end{pmatrix} \right)^{-1} \begin{pmatrix} \Phi^T & \sqrt{\lambda}I_{p \times M}^T \end{pmatrix} \begin{pmatrix} \mathbf{z} \\ \mathbf{0}_p \end{pmatrix} \\
	       &= \left( \lambda I_{M \times M} + \Phi^T\Phi \right)^{-1} \Phi^T \mathbf{z} \\
	       &= \mathbf{w}_{MLE}^{ridge}
\end{align}
%\end{equation}

\section*{Problem 2}

We want to prove the following identity:

\[ (M + vv^T)^{-1} = M^{-1} - \frac{(M^{-1} v) (v^T M^{-1})  }{1 + v^T M^{-1} v} \]

To prove the identity we will use the fact that for any matrix M:

\[
M M^{-1} = I
\]

where I is the identity matrix. So we multiply both sides by the inverse of the left-hand side

\[
M + vv^T
\]

which results in

\begin{equation}
I = I + M^{-1}vv^T - \frac{M^{-1} v v^T + M^{-1} v v^T M^{-1} v v^T}{1 + v^T M^{-1} v}
\end{equation}

$ v^T M^{-1} v $ is a scalar, so:

\[
M^{-1} v v^T + M^{-1} v v^T M^{-1} v v^T = ( 1 + v^T M^{-1} v ) M^{-1} v v^T
\]

(5) then simplifies to:

\[
I = I + M^{-1}vv^T - M^{-1}vv^T = I
\]

so the identity is verified.

Using the identity above, we need to prove the following inequality:

\begin{equation}
\sigma^2_{N+1} (x) \leq \sigma^2_{N} (x)
\end{equation}

We use the definition of $\sigma^2_{N}$ and the definition of $S_N$ (here $ \Phi_N $ is the design matrix with the first N observations).

\[
\sigma^2_{N} (x) = \frac{1}{\beta} + \phi(x)^T S_N \phi(x)
\]

\[
S^{-1}_{N} = \alpha I + \beta (\Phi^T_{N}) \Phi_N
\]

We now use the following identity:

\[
(\Phi^T_{N}) \Phi_N = \sum_{k=1}^N \phi(x_k) \phi(x_k)^T
\]

so we can write

\[
(\phi^T_{N+1}) \phi_{N+1} =(\phi^T_{N}) \phi_N + \phi(x_{N+1}) \phi(x_{N+1})^T
\]

and hence

\[
S^{-1}_{N+1} = \alpha I + \beta (\Phi^T_{N+1}) \Phi_{N+1} = \alpha I + \beta ((\Phi^T_{N}) \Phi_N + \phi(x_{N+1}) \phi(x_{N+1})^T) = S^{-1}_{N} + \phi(x_{N+1}) \phi(x_{N+1})^T
\]

Substituting this into $ \sigma^2_{N+1} (x) $ and using the above matrix identity we get:

\[
\sigma^2_{N+1} (x) = \frac{1}{\beta} + \phi(x)^T S_{N+1} \phi(x)
\]

\[
 = \frac{1}{\beta} + \phi(x)^T ({S^{-1}_{N} + \phi(x_{N+1}) \phi(x_{N+1})^T })^{-1} \phi(x)
\]

\[
 = \frac{1}{\beta} + \phi(x)^T (S_{N} - \frac{S_{N} \phi(x_{N+1}) \phi(x_{N+1})^T S_{N}} {1 + \phi(x_{N+1})^T S_{N} \phi(x_{N+1})}) \phi(x) \]

\begin{equation}
 = \frac{1}{\beta} + \phi(x)^T S_{N} \phi(x) - \frac{ \phi(x)^T S_{N} \phi(x_{N+1}) \phi(x_{N+1})^T S_{N} \phi(x)} {1 + \phi(x_{N+1})^T S_{N} \phi(x_{N+1})}
\end{equation}

The nominator of the last fraction is greather than zero. Now note that the $S^{-1}_N$ is positive semi-definite since $\beta$ is greater than zero, $\alpha$ is greater than zero, and:

\[
x^T S^{-1}_{N} x = x^T (\alpha I + \beta \Phi^{T}_{N} \Phi_N  ) x = x^T \alpha I x + x^T \beta \Phi^{T}_{N} \Phi_N x = \alpha x^T I x + \beta (\Phi_N x)^T \Phi_N x
\]

A matrix A has an eigenvalue if and only if $A^{-1}$ has eigenvalue $\frac{1}{\lambda}$, as

\[
Av = \lambda v \Leftrightarrow A^{-1} A v = \lambda A^{-1} v \Leftrightarrow \frac{1}{\lambda} v = A^{-1} v 
\]

With that and the fact that the eigenvalues of positive semi-definite matrices are nonnegative we have shown
that $S_N$ also is positive semi-definite. Therefore

\[
\phi(x_{N+1})^T S_{N} \phi(x_{N+1}) \geq 0
\]

and so the denominator of the last fraction in eq. (7) is positive. We thus have proven the inequality.

\section*{Problem 3}
Since we spent no money, the model boils down to $y = 10 + \mathcal{N}(0,4)$. The Gaussian is centered on zero and, therefore, we can say that the probability to get more than 10 likes is $0.5$.

\section*{Problem 4}

Spending 1 EUR on advertisements our model gives us 
\begin{equation}
y = 15 + \mathcal{N}(0,4) \hspace{0.5cm} \Rightarrow \hspace{0.5cm} \operatorname{E}[y] = 15
\end{equation}
with \hspace{0.5cm} $\operatorname{E}[\mathcal{N}(0,4)] = 0$. Hence, we expect 15 likes.

### 2015-12-02 ML/hw7/hw.tex
ml homework and some random previously unsubmitted files for paralle_programming and HiWi

>>>>

\section*{Problem 1}
If the data set is linearly separable, any decision boundary separating the two classes will have the property
\begin{equation}
w^T\phi(x^n) = \begin{cases} \geq 0 & \mbox{if } z^n = 1, \\ < 0 & otherwise. \\ \end{cases}
\label{eq:}
\end{equation}

Moreover, the log-likelihood 
\begin{align}
L(w) &= \log p(\{z^i\}|w,\{\phi(x^i)\})\\
&= \sum^{N}_{n = 1}\left[z^n\log \sigma(w^T\phi(x^n)) + (1-z^n)(\log(1-\sigma(w^T\phi(x^n))))\right]
\end{align}

will be maximized when $y^n = \sigma(w^T\phi(x^n)) = z^n \hspace{0.3cm}\forall n$. This will be the case when the sigmoid function is saturated, which occurs when its argument, $w^T\phi$, goes to $\pm \inf$, i.e., when the magnitude of $w$ goes to infinity.

\section*{Problem 2}
The basis function 
\begin{equation}
\phi(x_1,x_2) = x_1x_2
\end{equation}

enables us to linearly seperate the crosses from the circles. The hyperplane would go through the origin of the coordinate system.
